---
title: "Formulating experimental feeds"
author: Tormey Reimer
date: today
format: html
editor: source
---

# Introduction

```{r setup}
library(arrow)
library(conflicted)
library(dplyr)
library(fs)
library(furrr)
library(future)
library(ggplot2)
library(here)
library(magrittr)
library(purrr)
library(qs)
library(readxl)
library(sf)
library(stringr)
library(targets)
library(terra)
library(tictoc)
library(tidyr)
library(units)

conflicts_prefer(dplyr::select(), dplyr::filter(), .quiet = T)

here("00_model_functions.R") %>% source()
here("00_dirs.R") %>% source()
```

# Make sure incoming data is complete

You don't really need to do this again unless you make changes - all the important stuff is already saved.

```{r precheck}
#| code-summary: Get temperature data and formulated feeds
#| eval: false

here("01_extracting_temperatures.R") |> source()
here("02_formulating_feeds.R") |> source()
```

## Choose "reference" feed

The "reference" feed is somewhat arbitrary. It's the only feed run in sensitivity analysis, and it's the point of comparison in the main analysis.

```{r reference-feed}
feed_params <- file.path(output_species_data_path, "feed_params.qs") %>% qs::qread()
reference_feed <- feed_params[["plant_dominant"]]
```

# Set up for ananlyses

This section makes sure all the data necessary for the individual runs (sensitivity) and farm growth pipelines is in the correct format.

```{r load-farm-times}
#| code-summary: Get farm locations

farm_coords_file <- file.path(farms_locs_path, "farm_coords.parquet")
farm_starts_file <- file.path(output_farm_data_path, "farm_starts.qs")
farm_geometry_file <- file.path(output_farm_data_path, "farm_geometry.qs")

farm_starts <- if (!file.exists(farm_coords_fortargets_file)) {
    times_N <- c("t_start" = 121, "t_end" = 121+547, "dt" = 1)
    times_S <- c("t_start" = 274, "t_end" = 274+547, "dt" = 1)

    fc <- farm_coords_file %>% 
    read_parquet() %>% 
    mutate(t_start = case_when(lat > 0 ~ times_N['t_start'], TRUE ~ times_S['t_start']), 
            t_end = case_when(lat > 0 ~ times_N['t_end'], TRUE ~ times_S['t_end']),
            t_start = unname(t_start),
            t_end = unname(t_end))
    qsave(fc, farm_starts_file)
    fc
} else {
  qs::qread(farm_starts_file)
}

# Also save geometry for later
if (!file.exists(farm_geometry_file)) {
  file.path(farms_locs_path, "atlantic_salmon_locations_w_temps.qs") %>% 
    qread() %>% 
    dplyr::filter(day == "day_1") %>% 
    dplyr::select(farm_id, geometry, country) %>% 
    qsave(farm_geometry_file)
}
```

```{r load-farm-temperature-data}
#| code-summary: Get temperature timeseries data

farm_ts_file <- file.path(output_farm_data_path, "farm_ts_data.qs")

farm_ts_data <- if (!file.exists(farm_ts_file)) {
    farms_to_omit <- file.path(farms_locs_path, "atlantic_salmon_farms_to_omit.qs") %>% qread()
    farm_SST_data <- file.path(farm_sst_path, "farm_SST_extracted.parquet") %>% read_parquet()
    
    df <- farm_SST_data %>%
        rename(farm_ID = farm_id) %>% 
        filter(!farm_ID %in% farms_to_omit) %>% 
        select(c(farm_ID, day, temp_c)) %>%
        mutate(day = str_split_i(day, "day_", 2) %>% as.integer())
    qsave(df, farm_ts_file)
    df
} else {
  qs::qread(farm_ts_file)
}
```

```{r load-species-parameters}
#| code-summary: Get species-specific parameters

species_params_excel <- c(
    file = file.path(input_species_param_path, "Species.xlsx"), 
    sheet = "data"
    )
species_params_file <- file.path(output_species_data_path, "species_params.qs")

species_params <- if (!file.exists(species_params_file)) {
    df <- readxl::read_excel(path = species_params_excel["file"], sheet = species_params_excel["sheet"])
    vc <- df$Value
    names(vc) <- df$Quantity
    vc <- vc[!is.na(vc)]
    qsave(vc, species_params_file)
    vc
} else {
  qs::qread(species_params_file)
}
```

```{r load-population-parameters}
#| code-summary: Get population parameters

pop_params_excel <- c(file = file.path(input_species_param_path, "Population.xlsx"))
pop_params_file <- file.path(output_species_data_path, "pop_params.qs")

pop_params <- if (!file.exists(pop_params_file)) {
    df <- readxl::read_excel(path = pop_params_excel["file"])
    vc <- df$Value
    names(vc) <- df$Quantity
    vc <- vc[!is.na(vc)]
    qsave(vc, pop_params_file)
    vc
} else {
  qs::qread(pop_params_file)
}
```

## Calculate harvest size

Individual harvest sizes are calculated for each farm. 

```{r calculate-farm-harvest-size}
#| code-summary: Calculate individual harvest size for each farm 

farm_harvest_size_file <- file.path(output_farm_data_path, "farm_harvest_size.qs")

farm_harvest_size <- if (file.exists(farm_harvest_size_file)) {
    qs::qread(farm_harvest_size_file)
} else {
    hs <- purrr::map(farm_IDs, function(farm_id) {
        farm_times <- c(
            t_start = farm_starts$t_start[farm_starts$farm_ID == farm_id],
            t_end = farm_starts$t_end[farm_starts$farm_ID == farm_id],
            dt = 1
        )
        farm_temp <- farm_ts_data %>%
            filter(farm_ID == farm_id) %>%
            filter(day >= farm_times['t_start'], day <= farm_times['t_end']) %>%
            pull(temp_c)
        df <- fish_growth(
            pop_params = pop_params,
            species_params = species_params,
            water_temp = farm_temp,
            feed_params = reference_feed,
            times = farm_times,
            init_weight = species_params["meanW"],
            ingmax = species_params["meanImax"]
        )
        df %>%
            as.data.frame() %>%
            slice_tail(n = 1) %>%
            select(weight) %>%
            mutate(farm_ID = farm_id,
                    weight = units::set_units(weight, "g"))
        }) %>% 
            bind_rows()

    qsave(hs, farm_harvest_size_file)
    hs
}
```

# Choose a farms subset

Normally the analysis would be run for all farms, but this takes a long time. For now I'm going to choose a sample of farms to run going forward.

```{r farm-subset}
farm_IDs <- farm_ts_data %>% 
  pull(farm_ID) %>% 
  unique() %>% 
  sample(10)

farm_starts <- farm_starts %>% filter(farm_ID %in% farm_IDs)

farm_ts_data <- farm_ts_data %>% 
  filter(farm_ID %in% farm_IDs) %>% 
  merge(farm_starts, by = "farm_ID") %>% 
  filter(between(day, t_start, t_end))
farm_ts_data <- split(farm_ts_data, farm_ts_data$farm_ID)

farm_harvest_size <- farm_harvest_size %>% 
  filter(farm_ID %in% farm_IDs) %>% 
  pull(weight)
```

# Run sensitivity analysis

The sensitivity analysis investigates the overall impact of changing each parameter by $\pm10\%$.

```{r set-up-sensitivity-runs}
param_names_pop <- names(pop_params)
param_names_pop <- param_names_pop[!param_names_pop %in% c("nruns")]

param_names_spec <- names(species_params)
param_names_spec <- param_names_spec[!param_names_spec %in% c('betaprot', 'betalip', 'betacarb', 'fcr', 'CS')]

all_params <- c(pop_params, species_params)

factors <- c(0.9, 1, 1.1)
```

These next blocks run the sensitivity analyses. 

```{r run-sensitivity-spec-params}
sens_spec <- purrr::map(factors, function(fac) {
  purrr::map(param_names_spec, function(nm) {
    purrr::map2(farm_IDs, farm_ts_data, function(farm, ts) {
      adj_params <- all_params
      adj_params[nm] <- adj_params[nm] * fac

      sens <- uni_farm_growth(
        pop_params = adj_params,
        species_params = adj_params,
        water_temp = ts$temp_c,
        feed_params = reference_feed,
        times = c('t_start' = unique(ts$t_start), 't_end' = unique(ts$t_end), 'dt' = 1),
        N_pop = rep(1, nrow(ts))
      )

      data.frame(
        weight = last(sens[['weight_stat']][, 2][!is.na(sens[['weight_stat']][, 2])]),
        dw = mean(sens[['dw_stat']][, 2], na.rm = T),
        total_excr = sum(sens[['total_excr_stat']][, 2], na.rm = T),
        P_excr = sum(sens[['P_excr_stat']][, 2], na.rm = T),
        L_excr = sum(sens[['L_excr_stat']][, 2], na.rm = T),
        C_excr = sum(sens[['C_excr_stat']][, 2], na.rm = T),
        total_uneat = sum(sens[['total_uneat_stat']][, 2], na.rm = T),
        P_uneat = sum(sens[['P_uneat_stat']][, 2], na.rm = T),
        L_uneat = sum(sens[['L_uneat_stat']][, 2], na.rm = T),
        C_uneat = sum(sens[['C_uneat_stat']][, 2], na.rm = T),
        food_prov = sum(sens[['food_prov_stat']][, 2], na.rm = T),
        rel_feeding = mean(sens[['rel_feeding_stat']][, 2], na.rm = T),
        ing_act = sum(sens[['ing_act_stat']][, 2], na.rm = T),
        anab = mean(sens[['anab_stat']][, 2], na.rm = T),
        catab = mean(sens[['catab_stat']][, 2], na.rm = T),
        O2 = sum(sens[['O2_stat']][, 2], na.rm = T),
        NH4 = sum(sens[['NH4_stat']][, 2], na.rm = T),
        adj_param = as.factor(nm)
      ) %>%
        mutate(farm_ID = farm)
    }) %>%
      bind_rows() %>% 
      mutate(factor = fac)
  }) %>% 
    bind_rows()  
}) %>% 
  bind_rows()  
```

Note that the sensitivity of population parameters are run for every factor, nruns x (5000). This might take a long time (~20 minutes).

```{r run-sensitivity-pop-params}
plan("multisession", workers = 15)

sens_pop <- future_map(factors, function(fac) {
  future_map(param_names_pop, function(nm) {

    adj_params <- all_params
    adj_params[nm] <- adj_params[nm] * fac

    future_map(farm_IDs, farm_ts_data, function(farm, ts) {

      sens <- farm_growth(
        pop_params = adj_params,
        species_params = adj_params,
        water_temp = ts$temp_c,
        feed_params = reference_feed,
        times = c('t_start' = unique(ts$t_start), 't_end' = unique(ts$t_end), 'dt' = 1),
        N_pop = rep(1, nrow(ts)),
        nruns = adj_params["nruns"] # the only difference
      )

      data.frame(
        weight = last(sens[['weight_stat']][, 2][!is.na(sens[['weight_stat']][, 2])]),
        dw = mean(sens[['dw_stat']][, 2], na.rm = T),
        total_excr = sum(sens[['total_excr_stat']][, 2], na.rm = T),
        P_excr = sum(sens[['P_excr_stat']][, 2], na.rm = T),
        L_excr = sum(sens[['L_excr_stat']][, 2], na.rm = T),
        C_excr = sum(sens[['C_excr_stat']][, 2], na.rm = T),
        total_uneat = sum(sens[['total_uneat_stat']][, 2], na.rm = T),
        P_uneat = sum(sens[['P_uneat_stat']][, 2], na.rm = T),
        L_uneat = sum(sens[['L_uneat_stat']][, 2], na.rm = T),
        C_uneat = sum(sens[['C_uneat_stat']][, 2], na.rm = T),
        food_prov = sum(sens[['food_prov_stat']][, 2], na.rm = T),
        rel_feeding = mean(sens[['rel_feeding_stat']][, 2], na.rm = T),
        ing_act = sum(sens[['ing_act_stat']][, 2], na.rm = T),
        anab = mean(sens[['anab_stat']][, 2], na.rm = T),
        catab = mean(sens[['catab_stat']][, 2], na.rm = T),
        O2 = sum(sens[['O2_stat']][, 2], na.rm = T),
        NH4 = sum(sens[['NH4_stat']][, 2], na.rm = T),
        adj_param = as.factor(nm)
      ) %>%
        mutate(farm_ID = farm)

    }, .options = furrr_options(seed = T)) %>%
      bind_rows() %>% 
      mutate(factor = fac)

  }, .options = furrr_options(seed = T)) %>% 
    bind_rows()  
  
}, .options = furrr_options(seed = T)) %>% 
  bind_rows()  

plan("sequential")
```

```{r process-sensitivities}
stat_names <- colnames(sens_spec)[!colnames(sens_spec) %in% c("farm_ID", "factor", "adj_param")]

sens_spec <- sens_spec %>% 
  group_by(farm_ID, factor, adj_param) %>% 
  slice_tail(n = 1) %>% 
  pivot_longer(
    cols = all_of(stat_names), 
    names_to = "measure", 
    values_to = "value", 
    names_transform = list(measure = as.factor)
    ) %>% 
  pivot_wider(
    names_from = "factor", 
    values_from = "value", 
    names_prefix = "fact_"
    ) %>% 
  mutate(sensitivity = (fact_1.1-fact_0.9)/(0.2*fact_1)) %>% 
  group_by(adj_param, measure) %>%
  reframe(
    mean_sens = meanna(sensitivity),
    sd_sens = sdna(sensitivity)
    )

sens_pop <- sens_pop %>% 
  group_by(farm_ID, factor, adj_param) %>% 
  slice_tail(n = 1) %>% 
  pivot_longer(
    cols = all_of(stat_names), 
    names_to = "measure", 
    values_to = "value", 
    names_transform = list(measure = as.factor)
    ) %>% 
  pivot_wider(
    names_from = "factor", 
    values_from = "value", 
    names_prefix = "fact_"
    ) %>% 
  mutate(sensitivity = (fact_1.1-fact_0.9)/(0.2*fact_1)) %>% 
  group_by(adj_param, measure) %>%
  reframe(
    mean_sens = meanna(sensitivity),
    sd_sens = sdna(sensitivity)
    )

sens_results <- rbind(sens_spec, sens_pop)

sens_results_files <- file.path(output_sens_data_path, paste0("sens_results_", stat_names, ".qs"))
sens_results_figfiles <- file.path(output_sens_data_path, paste0("sens_plot_", stat_names, ".qs"))

for (sn in seq_along(stat_names)) {
    sr <- sens_results %>% 
        filter(measure == stat_names[sn]) 
    qsave(sr, sens_results_files[sn])

    p <- sr %>% 
        ggplot(aes(x = adj_param, y = mean_sens, ymin = mean_sens-sd_sens, ymax = mean_sens+sd_sens)) +
        geom_col(fill = "salmon", alpha = 0.65, colour = "black") +
        geom_errorbar(width = 0.5) +
        coord_flip() +
        theme_classic()
    qsave(p, sens_results_figfiles[sm])
}
```

# Run farm growth 

Much of the data needed for this has already been loaded earlier in this markdown.

```{r generate-population}
farm_static_data <- file.path(farms_locs_path, "atlantic_salmon_locations_w_temps.qs") %>% 
      qs::qread() %>% 
      sf::st_drop_geometry() %>% 
      distinct(farm_id, tonnes_per_farm) %>% 
      mutate(
        tonnes_per_farm = tonnes_per_farm %>% 
          units::set_units("t") %>% 
          units::set_units("g") %>% 
          units::drop_units()
        ) %>% 
      rename(farm_ID = farm_id) %>% 
      merge(farm_harvest_size, by = "farm_ID")

farm_static_data <- split(farm_static_data, farm_static_data$farm_id)

N_pop <- purrr::map2(farm_static_data, farm_starts, function(sd, fs) {
  generate_pop(
    harvest_n = sd$tonnes_per_farm/sd$harvest_size,
    mort = pop_params['mortmyt'],
    times = c(unique(fs['t_start']), unique(fs['t_end']), 'dt' = 1)
    )
}
```

Like the sensitivity pipeline, this takes a long time to run (4-5 minutes per iteration).

```{r run-farm-growth-pipeline}
tar_make(names = c("tar_harvest_size", "tar_common_params", "tar_N_pop", "tar_farm_ts"), seconds_meta_append = 90)
```

```{r tar-farmruns-comparisons}
patt <- tar_pattern(
    cross(fi, fn),
    fi = length(farm_IDs),
    fn = length(feed_names)
) %>% 
    mutate(
        fi = factor(fi, labels = farm_IDs) %>% as.character() %>% as.integer(),
        fn = factor(fn, labels = feed_names)
    )
patt$branch <- 1:nrow(patt)

stat_names <- c(
    "weight_stat", 
    "dw_stat", 
    "water_temp_stat", 
    "T_response_stat", 
    "P_excr_stat", 
    "L_excr_stat", 
    "C_excr_stat", 
    "P_uneat_stat", 
    "L_uneat_stat", 
    "C_uneat_stat", 
    "food_prov_stat", 
    "food_enc_stat", 
    "rel_feeding_stat", 
    "ing_pot_stat", 
    "ing_act_stat", 
    "E_assim_stat", 
    "E_somat_stat", 
    "anab_stat","catab_stat", 
    "O2_stat", 
    "NH4_stat", 
    "total_excr_stat", 
    "total_uneat_stat", 
    "metab_stat",
    "biomass_stat"
    )

# Convert farms to cohorts (reference feed)
tar_farmrun_comparisons <- purrr::map(feed_names, function(fn) {
    tar_load(tar_farmrun, branches = patt$branch[patt$fi == 215 & patt$fn == fn])
    ref1 <- lapply(tar_farmrun, farm_to_cohort, time_offset = 0)
    ref2 <- lapply(tar_farmrun, farm_to_cohort, time_offset = 365)
    ref3 <- lapply(tar_farmrun, farm_to_cohort, time_offset = 730)
    lims <- unique(ref2[[1]]$t)

    ref <- purrr::map(1:length(stat_names), function(st) {
        df <- bind_rows(
            ref1[[st]] %>% mutate(cohort = 1), 
            ref2[[st]] %>% mutate(cohort = 2), 
            ref3[[st]] %>% mutate(cohort = 3)
            ) %>% 
          mutate(sd = sd/mean, 
                 feed = factor(fn, levels = feed_names)) %>%
          group_by(feed, t) %>%
          reframe(mean = sum(mean),
                  sd = sum(sd)*mean,
                  stat = stat_names[st]) %>% 
          filter(t %in% lims)
      })
}) %>% 
    bind_rows() %>% 
    mutate(
        farm_ID = as.integer(73),
        t = as.integer(t)
        )




tar_target(
    tar_farmrun_comparisons,
    command = {
      # Convert farms to cohorts (reference feed)
      ref1 <- lapply(tar_farmrun_reference, farm_to_cohort, time_offset = 0)
      ref2 <- lapply(tar_farmrun_reference, farm_to_cohort, time_offset = 365)
      ref3 <- lapply(tar_farmrun_reference, farm_to_cohort, time_offset = 730)
      
      # Get the t values of cohort 2
      lims <- unique(ref2[[1]]$t)
      
      ref <- purrr::map(1:length(stat_names), function(st) {
        bind_rows(ref1[[st]], ref2[[st]], ref3[[st]]) %>% 
          mutate(sd = sd/mean, 
                 feed = factor("reference", levels = c("reference", "past", "future"))) %>%
          group_by(feed, t) %>%
          reframe(mean = sum(mean),
                  sd = sum(sd)*mean) %>% 
          filter(t %in% lims)
      })
      
      # Convert farms to cohorts (future feed)
      ref1 <- lapply(tar_farmrun_future, farm_to_cohort, time_offset = 0)
      ref2 <- lapply(tar_farmrun_future, farm_to_cohort, time_offset = 365)
      ref3 <- lapply(tar_farmrun_future, farm_to_cohort, time_offset = 730)
      
      # Get the t values of cohort 2
      lims <- unique(ref2[[1]]$t)
      
      fut <- purrr::map(1:length(stat_names), function(st) {
        bind_rows(ref1[[st]], ref2[[st]], ref3[[st]]) %>% 
          mutate(sd = sd/mean, 
                 feed = factor("future", levels = c("reference", "past", "future"))) %>%
          group_by(feed, t) %>%
          reframe(mean = sum(mean),
                  sd = sum(sd)*mean) %>% 
          filter(t %in% lims)
      })
      
      # Convert farms to cohorts (past feed)
      ref1 <- lapply(tar_farmrun_past, farm_to_cohort, time_offset = 0)
      ref2 <- lapply(tar_farmrun_past, farm_to_cohort, time_offset = 365)
      ref3 <- lapply(tar_farmrun_past, farm_to_cohort, time_offset = 730)
      
      # Get the t values of cohort 2
      lims <- unique(ref2[[1]]$t)
      
      pas <- purrr::map(1:length(stat_names), function(st) {
        bind_rows(ref1[[st]], ref2[[st]], ref3[[st]]) %>% 
          mutate(sd = sd/mean, 
                 feed = factor("past", levels = c("reference", "past", "future"))) %>%
          group_by(feed, t) %>%
          reframe(mean = sum(mean),
                  sd = sum(sd)*mean) %>% 
          filter(t %in% lims)
      })
      
      # Put them all together
      purrr::map(1:length(stat_names), function(st) {
        bind_rows(ref[[st]], pas[[st]], fut[[st]]) %>% 
          mutate(farm_ID = as.integer(tar_farm_IDs),
                 t = as.integer(t))
      })
    },
    pattern = map(tar_farm_IDs, tar_farmrun_reference, tar_farmrun_past, tar_farmrun_future)
  )
```

```{r save-targets-outputs-2}
tar_load(stat_names)

farmrun_reference_files <- file.path(output_growth_data_path, sprintf("farmrun_reference_%s.qs", fixnum(farm_IDs,4)))
farmrun_past_files <- file.path(output_growth_data_path, sprintf("farmrun_past_%s.qs", fixnum(farm_IDs,4)))
farmrun_future_files <- file.path(output_growth_data_path, sprintf("farmrun_future_%s.qs", fixnum(farm_IDs,4)))
farmrun_comparisons_files <- file.path(output_cohorts_data_path, sprintf("farmrun_comparisons_%s.qs", fixnum(farm_IDs,4)))


for (f in seq_along(farm_IDs)) {
  targets::tar_load(tar_farmrun_reference, branches = f) 
  tar_farmrun_reference %>%
    setNames(stat_names) %>%
    lapply(function(x) {colnames(x) <- c("t", "mean", "sd"); x}) %>%
    lapply(function(x) {cbind(x, farm_ID = farm_IDs[f])})  %>%
    qs::qsave(farmrun_reference_files[f])
  
  targets::tar_load(tar_farmrun_past, branches = f) 
  tar_farmrun_past %>%
    setNames(stat_names) %>%
    lapply(function(x) {colnames(x) <- c("t", "mean", "sd"); x}) %>%
    lapply(function(x) {cbind(x, farm_ID = farm_IDs[f])})  %>%
    qs::qsave(farmrun_past_files[f])
  
  targets::tar_load(tar_farmrun_future, branches = f) 
  tar_farmrun_future %>%
    setNames(stat_names) %>%
    lapply(function(x) {colnames(x) <- c("t", "mean", "sd"); x}) %>%
    lapply(function(x) {cbind(x, farm_ID = farm_IDs[f])})  %>%
    qs::qsave(farmrun_future_files[f])
  
  # Save the final cohort feed comparison data
  targets::tar_load(tar_farmrun_comparisons, branches = f) 
  tar_farmrun_comparisons %>% 
    setNames(stat_names) %>% 
    qs::qsave(farmrun_comparisons_files[f])
}

# Use the intermediate growth data to save aggregate files
aggregate_comparison_files <- file.path(output_cohorts_data_path, str_c("allfarms_comparisons_", stat_names, ".qs"))

for (s in seq_along(stat_names)) {
  purrr::map(farmrun_comparisons_files, function(fnm) {
    qread(fnm)[[stat_names[s]]]
  }) %>% 
    bind_rows() %>% 
    qsave(aggregate_comparison_files[s])
}
```




